


```{r}
library(tidyverse)
library(mice)
library(lattice)
```



## finding the average NSAF values of proteins in controlled group
```{r}
protein_id <- read.csv("protein_id.csv")
protein_id
ctrl_group <-protein_id[grep("ctrl", protein_id$sample),]
ctrl_group

# group by time points
# day 0
ctrl_day0 <-ctrl_group[seq(1,36,by=7),] %>% select(-c(X, sample))
ctrl_day0
# day 1
ctrl_day1 <-ctrl_group[seq(2,37,by=7),] %>% select(-c(X, sample))
# day 3
ctrl_day3 <-ctrl_group[seq(5,40,by=7),] %>% select(-c(X, sample))
# day 5
ctrl_day5 <-ctrl_group[seq(6,41,by=7),] %>% select(-c(X, sample))
# day 7
ctrl_day7 <-ctrl_group[seq(7,42,by=7),] %>% select(-c(X, sample))
# day 10
ctrl_day10 <-ctrl_group[seq(3,38,by=7),] %>% select(-c(X, sample))
# day 14
ctrl_day14 <-ctrl_group[seq(4,39,by=7),] %>% select(-c(X, sample))


avg_protein_ctrl <- cbind(as.data.frame(colMeans(ctrl_day0)), as.data.frame(colMeans(ctrl_day1)),as.data.frame(colMeans(ctrl_day3)),as.data.frame(colMeans(ctrl_day5)),as.data.frame(colMeans(ctrl_day7)),as.data.frame(colMeans(ctrl_day10)),as.data.frame(colMeans(ctrl_day14)))
names(avg_protein_ctrl) <- c("day_0","day_1","day_3","day_5","day_7","day_10","day_14")
avg_protein_ctrl


sum(is.na(avg_protein_ctrl))/(dim(avg_protein_ctrl)[1]*dim(avg_protein_ctrl)[2])*100

write.csv(avg_protein_ctrl,file="protein_control.csv")
```



strains:
aj, balbc, c57, cej, dba, fvb

keep the ones with at least 4 time values in both files (if >1000)
look for methods in r that can 
replace NA values with mean
calculate variance for each protein
average variance of all proteins

# filter out rows with >4 NA

```{r}
na_count_ctrl <- as.data.frame(rowSums(is.na(avg_protein_ctrl)))
colnames(na_count_ctrl) <- "na_count"
na_count_ctrl
# get rows with at least 4 values
filtered_ctrl <-avg_protein_ctrl[which(na_count_ctrl$na_count < 4),]
filtered_ctrl
# get average of each row
row_mean <-rowMeans(filtered_ctrl,na.rm=TRUE)

# replace NA with row average
# loop through row
for(i in 1:dim(filtered_ctrl)[1]){
  # loop through col
  for(j in 1:dim(filtered_ctrl)[2])
    if(is.na(filtered_ctrl[i,j]) == TRUE){
      filtered_ctrl[i,j] <- row_mean[i]
    }
}

filtered_ctrl

# variance of filtered_ctrl dataset: 0.0478332
sum(apply(filtered_ctrl,1,var))/(dim(filtered_ctrl)[1])

```




```{r}
avg_protein_ctrl
# impute missing data
imputed_data_ctrl <- mice(avg_protein_ctrl, m=5, maxit = 10, method = 'pmm', seed = 500)
summary(imputed_data_ctrl)

imputed_data_ctrl
# save the 5 datasets with imputed values
mice_ctrl_1 <- mice::complete(imputed_data_ctrl,1)
mice_ctrl_2 <- mice::complete(imputed_data_ctrl,2)
mice_ctrl_3 <- mice::complete(imputed_data_ctrl,3)
mice_ctrl_4 <- mice::complete(imputed_data_ctrl,4)
mice_ctrl_5 <- mice::complete(imputed_data_ctrl,5)

write.csv(mice_ctrl_1, "mice_ctrl_1.csv")
write.csv(mice_ctrl_2, "mice_ctrl_2.csv")
write.csv(mice_ctrl_3, "mice_ctrl_3.csv")
write.csv(mice_ctrl_4, "mice_ctrl_4.csv")
write.csv(mice_ctrl_5, "mice_ctrl_5.csv")

mice_ctrl_1

# variance of mice_ctrl_1 dataset: 0.07275378
sum(apply(mice_ctrl_1,1,var))/dim(mice_ctrl_1)[1]
sum(apply(mice_ctrl_1,1,var))
```




multiple imputation
impute multiple values (5 datasets)
treating real values differently from imputed values


consider proteins in both contrl and iso
merge into original shape


still keep treatment col
strain/time/group

go back to original dataset and remove the averaging strain value step
select proteins present in 1) ctrl & iso 2) at least 3 and more
don't separate by treatment because it's already separated for us

so in mice, use logistic and poly for method to to imputation
still copy day0 from control for iso


col: group strain time points(7)

for mice:
group: logistic
strain: poly
time points: pmm





